Analysis of observational data, to calculate the Delayed-Tau model parameters

### 2025-02-11 thesis.qmd
small changes

  @haslbauerCosmologicalStarFormation2023 argue that the Delayed-$\tau$ model is opposed to the Lilly-Madau plot ([@madauCosmicStarFormation2014]), which plots the observed SFR's of galaxies with the corresponding redshifts ($z$) and calculates a cosmic SFR peak at $z\approx 2$. The way they calculated this inconsistency is by using observatory data for SFR and Stellar Masses from the UNGC catalog (@karachentsevUPDATEDNEARBYGALAXY2013, @karachentsevSTARFORMATIONPROPERTIES2013) for calculating the parameters (the timescale $\tau$ and the normalization constant $A_{del}$) of the model. This calculation for the galaxies of the Local Cosmological Volume (LV), allows the investigation of the SFR throughout the life each galaxy and so we can find the expected time of peak of the SFR.

@@

-   **Elliptical Galaxies (E):** Ellipsoidal shapes with smooth brightness profiles, containing older stars and minimal interstellar matter. Their eccentricity takes values from 0 to 0.7, so we can categorize them even further, form E0 to E7

@@

-   **H**\boldmath$\alpha$ **Emission**: Young, hot, massive stars (O-type stars, \~10 Myr, \~20 $M_\odot$) produce a number of ionizing photons, which they ionize the surrounding hydrogen rich gas. The hydorgen undergoes recombination cascades which produce Balmer emission lines of $H\alpha$ (0.6563 $\mu m$) and $H\beta$ (0.4861 $\mu m$). Dust can significantly affect observations.

@@

But, because light takes time to cover the distance $d$ between two galaxies, when the light finally reaches as, we will see the observed galaxy, as it was when the light was emitted, and not how it is at this moment. If we substitute time that it took the light to reach us over the distance, then we arrive at the relation [@longairGalaxyFormation1998]:

@@

The Newton-Raphson method is one of the most common numerical methods for estimating the root of a real-valued function. This method is based on the concept of linear approximation, where the function is locally approximated by its tangent line. It starts with initial guess of the root and iteratively refines the result using a formula that involves derivative of the function.

Let $x_n$ be the n-th approximation of the root and $f(x), f'(x)$ the function and the derivative, respectivelly. Then the refined approximation of the root $x_{n+1}$ is given by the formula

@@

A great example of such a chain is the board game snake and ladders. The dice has a uniform probabiliy distribution across 6 stages (integers 1 to 6). The players have a position on the board, but their next position on the board is only based on the current position and the random roll of the dice. The one dimension random walk algorithm is also a Markov Chain.

@@

Let $x^{(i)}$ be a homoguneous Markov chain[^5] inside a high-dimensional space $X$ of the set of possible configurations of a system and the space on which the posterior is defined, $x^{(i)}\in X = {x_1,x_2,...}$.

@@

In practice, this means that given a prior distribution for the parameters of the examed model (in our case we assume priors for the parameters $t_{sf},\ \tau,\ A_{del}$ from the delayed-$\tau$ model) and the likelihood based on observed data (we have experimental data for the SFR and $M_*$), a MCMC algorithm can calculate the posterior probability of the model and thus approximate the distribution of the parameters. From these distributions, their mean values are the values of the parameters and their variance is the error.

@@

We will join the two catalogs, based on the coordinates of the galaxies, so we could assure a bigger and more coplete sample size for our analysis. But first we need to undestand them and examine if there are any inconsistances between UNGC and HECATE.

>>>>

  @haslbauerCosmologicalStarFormation2023 argue that the Delayed-$\tau$ model is opposed to the Lilly-Madau plot ([@madauCosmicStarFormation2014]), which plots the observed SFR's of galaxies with the corresponding redshifts ($z$) and calculates a cosmic SFR peak at $z\approx 2$. The way they calculated this inconsistency is by using observatory data for SFR and Stellar Masses from the UNGC catalog (@karachentsevUPDATEDNEARBYGALAXY2013, @karachentsevSTARFORMATIONPROPERTIES2013) for calculating the parameters (the timescale $\tau$ and the normalization constant $A_{del}$) of the model. This calculation for the galaxies of the Local Cosmological Volume (LV), allows the investigation of the SFR throughout the life of each galaxy and so we can find the expected time of peak of the SFR.

@@

-   **Elliptical Galaxies (E):** Ellipsoidal shapes with smooth brightness profiles, containing older stars and minimal interstellar matter. Their eccentricity takes values from 0 to 0.7, so we can categorize them even further, from E0 to E7

@@

-   **H**\boldmath$\alpha$ **Emission**: Young, hot, massive stars (O-type stars, \~10 Myr, \~20 $M_\odot$) produce a number of ionizing photons, which they ionize the surrounding hydrogen rich gas. The hydrogen undergoes recombination cascades which produce Balmer emission lines of $H\alpha$ (0.6563 $\mu m$) and $H\beta$ (0.4861 $\mu m$). Dust can significantly affect observations.

@@

But, because light takes time to cover the distance $d$ between two galaxies, when the light finally reaches us, we will see the observed galaxy, as it was when the light was emitted, and not how it is at this moment. If we substitute time that it took the light to reach us over the distance, then we arrive at the relation [@longairGalaxyFormation1998]:

@@

The Newton-Raphson method is one of the most common numerical methods for estimating the root of a real-valued function. This method is based on the concept of linear approximation, where the function is locally approximated by its tangent line. It starts with an initial guess of the root and iteratively refines the result using a formula that involves derivative of the function.

Let $x_n$ be the n-th approximation of the root and $f(x), f'(x)$ the function and the derivative, respectively. Then the refined approximation of the root $x_{n+1}$ is given by the formula

@@

A great example of such a chain is the board game "Snake and Ladders". The dice has a uniform probability distribution across 6 stages (integers 1 to 6). The players have a position on the board, but their next position on the board is only based on the current position and the random roll of the dice. The one dimension random walk algorithm is also a Markov Chain.

@@

Let $x^{(i)}$ be a homogeneous Markov chain[^5] inside a high-dimensional space $X$ of the set of possible configurations of a system and the space on which the posterior is defined, $x^{(i)}\in X = {x_1,x_2,...}$.

@@

In practice, this means that given a prior distribution for the parameters of the examined model (in our case we assume priors for the parameters $t_{sf},\ \tau,\ A_{del}$ from the delayed-$\tau$ model) and the likelihood based on observed data (we have experimental data for the SFR and $M_*$), a MCMC algorithm can calculate the posterior probability of the model and thus approximate the distribution of the parameters. From these distributions, their mean values are the values of the parameters and their variance is the error.

@@

We will join the two catalogs, based on the coordinates of the galaxies, so we could assure a bigger and more complete sample size for our analysis. But first we need to undestand them and examine if there are any inconsistances between UNGC and HECATE.
